## Artificial Intelligence in Digital Archaeology 

### Agent-based modeling (ABM)

>Andreas: "Hello World!"
Overview

#### Models, simulation, and the precursors of ABM

>"Essentially, all models are wrong, but some are useful"  
>George Box, 1987  
>*Empirical Model-Building and Response Surfaces*, p. 424

- def system, model, formal model
- The origins of computer simulation
- Equation-based modeling, System Dynamics - ODE systems as models (mention Lorenz attractor, end with the idea of iteration of a process)
- Algorithm-based modeling, Algorithms as behavioral models (mention cellular automata, Van Neumann, Schelling, etc)
- The variety of algorithm-based models, much coincides with the variety of programming paradigm.
a. Flow: spectrum between centralized (iteration of a single process, e.g. event) and distributed (multiple processes are iterated contextually by multiple entities).
b. Schedule: spectrum between fully-scheduled (processes are initiated in a predetermined order) and event-driven (processes are initiated given a certain input).
ABMs fall somewhere in-between these two spectra, though normally leaning towards distributed (e.g., using Monte Carlo methods) and fully-scheduled. Applications in several disciplines (see Wikipedia)

#### ABM in History and Social Sciences

- Emergence... Life game, deterministic chaos
- Saving the gap between atomism and holism, or individual and social structure
- Virtual laboratory for social sciences
- Examples
- Mention the cycles of enthusiasm and skepticism, and the tendency to develop closed circles
- Challenges and pitfalls 

### discussion

#### Why model with ABM?

- open with the idea that we are always modeling
- Mention review article 'Why model?'

#### How to model with ABM?

- Parts of an ABM model
 (re-visit presentations)
- Define hypotheses, questions or simply the phenomenon of interest (identify the system)
- Define the elements and processes (you believe are) required to address the system (model the system). mention paradox Occam's razor vs. emergence
- Express the chosen elements and processes as computer code (implement the model)
- Modeling is a process of constant iteration. Each stage is an iteration loop that is often repeated several times before jumping into the next step.  The end result of development and exploration of a model may be relatively stable, often in the form of publications. However, this product potentially feeds a new  modeling effort (and that is actually the main usefulness of models in the long run). 
- ABM in python, Java, C#, or C++? in R? in NetLogo? (Any language can do it, really). Pros and cons, talk in terms of higher/lower level programming languages (still, they are all quite high-level).
- remember to talk about https://www.openabm.org/

### exercises

- The Pond Trade (coastal settlements of variable size around a pond; cultural vectors; trade ships travel between the settlements, re-evaluating afterwards if the trip is worse or better than other previous trips (make table); they carry economic value and cultural traits from between the base and destination settlements)

```{r, echo = FALSE}
library(DiagrammeR)
grViz("diagrams/boxes.dot")
```

- Implement the model based on general definition (agents, variables, parameters)
- Walk-through the implementation in NetLogo
- Simulations and experiment design
- Analysis and display of results (R example)

```{r, echo = FALSE}
plot(1:10, 1:10)
```

- Next steps

### Machine learning for image captioning and other classificatory tasks

blah

### discussion

blah

### exercises